Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-10-09 main_REV.tex

	% \end{subfigure} \hfill
  %   %
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	% \end{subfigure} \hfill
	\caption{
Three-dimensional rendering of the time evolution of
an initially orthogonal vortex configuration 
undergoing a vortex reconnection at $T = 1.9$K at dimensionless
times (from left to right)
$(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$, where $\tau_R = L_R/E^{1/2}_R$. The quantities $E_R$ and $L_R$ are the energy $E$ and integral length scales $L$ of the normal fluid at reconnection time $t_R$. The green
tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R=8.79\times10^{-4}$ . In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
    \label{fig:visualisation}
\end{figure*}

@@

The evolution of the vortex reconnection of a single pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by mutual friction, here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2025experimental}. The second row shows the rendering of the local helicity $H(\x)=\v_n\cdot\bom$, where we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during reconnection there is a net 
sudden normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity}. We will come back to this finding later.
\begin{figure}[t]
    \centering
    \includegraphics*[width=0.48\textwidth]{helicity_total.pdf}
\caption{Temporal evolution of the normal fluid helicity $\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$. For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$. $\tau_R = L_R/E^{1/2}_R$.
%\red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the two reconnections occurring at different times; same colour scheme as in Fig.~\ref{fig:visualisation}.}
}
\label{fig:total-helicity}
\end{figure}

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection. The energy spectrum $E(k)$ is displayed in Fig.~3a.
\begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]{energy-flux-spec.pdf}

    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{energy-spec.pdf}
    %     \caption{}\label{fig:kinetic-energy}
    % \end{subfigure}
    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{flux-spec.pdf}
    %     \caption{}\label{fig:energy-flux}
    % \end{subfigure}
% \begin{subcaptiongroup*}
% 	\phantomsubcaption\label{fig:minimum-distance}
% 	\phantomsubcaption\label{fig:prefactors}
% \end{subcaptiongroup*}

    \caption{\emph{(a):} Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines) for $T=1.9K$ (red)
and $T=2.1K$ (blue).\emph{(b):} Spectral normal fluid kinetic energy flux, $\Pi (k)$. It is normalised using the integral scale and the normal fluid energy at reconnection.
\emph{Inset:} Post reconnection evolution of the integral length scale, $L$.
Times and temperatures are labelled as in Fig.~3a.}
\label{fig:spectra}
\end{figure}
%

@@

% the dynamics of the normal fluid flow using the 
% spectral energy budget equation: 
% %
% \begin{equation}
%     \frac{\partial E(k)}{\partial t} = T(k) - D(k) + I(k)
% \end{equation}
% where $T(k)$ is the spectral kinetic energy transfer function,
% %(\textit{i.e.} $T(k)dk$ is the energy transfered per unit time by modes with $k < |\mathbf{k}| < k +dk$, determined by non-linear effects), 
% $D(k)=2\nu_n k^2 E_k$ is the dissipation spectrum and
% %=\mathrm{Re}(\hat{\mathbf{F}}_{ns}(\k)\cdot\hat{\v}^*(\k))$ 
% $I(k)$ is the injection spectrum arising from the mutual friction 
% force $\mathbf{F}_{ns}$.  
We observe that $\Pi(k) < 0$ for all $k$ during and after reconnection;
we also observe that, near the time of reconnection,
the peak value of $|\Pi(k)|$ is in the range $15 < k < 25$.
%
The negative sign of $\Pi(k)$ is evidence of a flux of kinetic  energy from small to large scales, exciting larger and larger scales. % In other words, at non-zero temperatures,
% vortex reconnections trigger an inverse transfer of energy, 
% which implies the creation of large scale structures, visible
% in Fig.~\ref{fig:visualisation}} \NOTE{GK: the figure shows small scale structures...}
This behaviour is quantified by the evolution of the integral length scale $L$, defined as
\begin{equation}
    L = \frac{\pi}{2 E}\int_0^{\infty}\frac{E(k)}{k}dk, %\Big/ \int_0^{\infty}\!\!E(k)dk.
\end{equation}
The inset of Fig.~3b shows that
$L$ indeed  increases steadily in the post-reconnection regime. Note that times have been normalised by the largest eddy-turnover-time at the reconnection event, evidencing  it fast evolution.

>>>>

	% \end{subfigure} \hfill
  %   %
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	% \end{subfigure} \hfill
	\caption{
Three-dimensional rendering of the time evolution of
an initially orthogonal vortex configuration 
undergoing a vortex reconnection at $T = 1.9$K at dimensionless
times (from left to right)
$(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$, where $\tau_R = L_R/E^{1/2}_R$. The green
tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R=8.79\times10^{-4}$ . In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
    \label{fig:visualisation}
\end{figure*}

@@

The evolution of the vortex reconnection of a single pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by mutual friction, here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2025experimental}. The second row shows the rendering of the local helicity $H(\x)=\v_n\cdot\bom$, where we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during reconnection there is a net 
sudden normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity}. We will come back to this finding later.
\begin{figure}[t]
    \centering

      \includegraphics[width=0.49\textwidth]{helicity_integralScale.pdf}
    %   \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{helicity_total.pdf}
    %     \caption{}\label{fig:}
    % \end{subfigure}
    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{integral-scale-scaled.pdf}
    %     \caption{}\label{fig:}

    % \end{subfigure}
\caption{\red{ \emph(a) Temporal evolution of the normal fluid helicity $\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$. \emph(b) Post reconnection evolution of the integral length scale, $L_0$ scaled by the reconnection time integral scale $L_R$. The bold solid lines represent the orthogonal reconnections (right axis), while the slightly faded lines represent the data at the same temperature for Hopf-link reconnections (left axis) from Ref.~\cite{stasiak2025experimental}.
%For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$.
}
%\red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the two reconnections occurring at different times; same colour scheme as in Fig.~\ref{fig:visualisation}.}
                       }
\label{fig:total-helicity}
\end{figure}

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection. The energy spectrum $E(k)$ is displayed in Fig.~3a.
\begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]{energySpec-fluxSpec.pdf}

    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{energy-spec-scaled.pdf}
    %     \caption{}\label{fig:kinetic-energy}
    % \end{subfigure}
    % \begin{subfigure}{0.49\textwidth}
    %     \centering
    %     \includegraphics*[width=\textwidth]{flux-spec-scaled.pdf}
    %     \caption{}\label{fig:energy-flux}
    % \end{subfigure}
\begin{subcaptiongroup*}
	\phantomsubcaption\label{fig:energy-spec}
	\phantomsubcaption\label{fig:flux-spec}
\end{subcaptiongroup*}

    \caption{\emph{(a):} Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines) for $T=1.9K$ (red)
and $T=2.1K$ (blue).\emph{(b):} Spectral normal fluid kinetic energy flux, $\Pi (k)$. \red{It is normalised using the box size and the quantum of circulation $\kappa$}. Times and temperatures are labelled as in Fig.~3a.}
\label{fig:spectra}
\end{figure}
%

@@

% the dynamics of the normal fluid flow using the 
% spectral energy budget equation: 
% %
% \begin{equation}
%     \frac{\partial E(k)}{\partial t} = T(k) - D(k) + I(k)
% \end{equation}
% where $T(k)$ is the spectral kinetic energy transfer function,
% %(\textit{i.e.} $T(k)dk$ is the energy transfered per unit time by modes with $k < |\mathbf{k}| < k +dk$, determined by non-linear effects), 
% $D(k)=2\nu_n k^2 E_k$ is the dissipation spectrum and
% %=\mathrm{Re}(\hat{\mathbf{F}}_{ns}(\k)\cdot\hat{\v}^*(\k))$ 
% $I(k)$ is the injection spectrum arising from the mutual friction 
% force $\mathbf{F}_{ns}$.  
We observe that $\Pi(k) < 0$ for all $k$ during and after reconnection;
we also observe that, near the time of reconnection,
the peak value of $|\Pi(k)|$ is in the range $15 < k < 25$.
%
The negative sign of $\Pi(k)$ is evidence of a flux of kinetic  energy from small to large scales, exciting larger and larger scales. % In other words, at non-zero temperatures,
% vortex reconnections trigger an inverse transfer of energy, 
% which implies the creation of large scale structures, visible
% in Fig.~\ref{fig:visualisation}} \NOTE{GK: the figure shows small scale structures...}
This behaviour is quantified by the evolution of the integral length scale $L_0$, defined as
\begin{equation}
    L_0 = \frac{\pi}{2 E}\int_0^{\infty}\frac{E(k)}{k}dk, %\Big/ \int_0^{\infty}\!\!E(k)dk.
\end{equation}
The inset of Fig.~3b shows that
$L_R$ indeed  increases steadily in the post-reconnection regime. Note that times have been normalised by the largest eddy-turnover-time at the reconnection event, evidencing  it fast evolution.
